\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-19 paper.qmd
Updated Figure 15

error_pbmc_umap <- read_rds("data/pbmc3k/error_pbmc_umap_30_min_dist_0.3.rds")
error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
# error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds") 
# error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds") 
# error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")

error_pbmc <- bind_rows(error_pbmc_umap, 
                        error_pbmc_umap2,
                        error_pbmc_umap3,
                        error_pbmc_tsne,
                        error_pbmc_tsne2)

@@

error_plot_pbmc <- ggplot(error_pbmc, 
                          aes(x = b_non_empty, 
                              y = Error/1000, 
                              colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  geom_vline(xintercept = 136, linetype="solid", 
             color = "black", linewidth=0.8, alpha = 0.5) +
  scale_color_manual(values=c('#ff7f00', '#984ea3', 
                              '#999999', '#a65628', '#f781bf')) +
  ylab("absolute error (in thousands)") +
  xlab("number of non-empty bins") + 
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7, angle = 90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        legend.position = "none") +
  inset_element(p = pbmc_layouts, 
                left = 1.55, 
                bottom = 0.35, 
                right = 0, 
                top = 1)

@@

```{r}
#| fig-cap: "Absolute error from UMAP and tSNE applied to PBMC3k dataset with diffferent (hyper-)parameter choices. What is the best (hyper-)parameter choice to create the model? The residual plot have a steep slope at the beginning, indicating that a smaller number of non-empty bins causes a larger amount of error. Then, the slope gradually declines or level off, indicating that a higher number of non-empty bins generates a smaller error. Using the elbow method, it was observed that when the number of non-empty bins is set to $136$, the lowest error occurred with the parameters perplexity: $30$."
#| label: fig-pbmc-abserror
#| fig-pos: H
#| out-width: 70%

>>>>

error_pbmc_umap <- read_rds("data/pbmc3k/error_pbmc_umap_30_min_dist_0.3.rds")
error_pbmc_umap2 <- read_rds("data/pbmc3k/error_pbmc_umap_5_min_dist_0.01.rds")
error_pbmc_umap3 <- read_rds("data/pbmc3k/error_pbmc_umap_15_min_dist_0.99.rds")
#error_pbmc_tsne <- read_rds("data/pbmc3k/error_pbmc_tsne_5.rds")
error_pbmc_tsne2 <- read_rds("data/pbmc3k/error_pbmc_tsne_30.rds")
# error_pbmc_phate <- read_rds("data/pbmc3k/error_pbmc_phate_5.rds") 
# error_pbmc_trimap <- read_rds("data/pbmc3k/error_pbmc_trimap_12_4_3.rds") 
# error_pbmc_pacmap <- read_rds("data/pbmc3k/error_pbmc_pacmap_30_random_0.9_5.rds")

error_pbmc <- bind_rows(error_pbmc_umap, 
                        error_pbmc_umap2,
                        error_pbmc_umap3,
                        error_pbmc_tsne2)

@@

error_plot_pbmc <- ggplot(error_pbmc, 
                          aes(x = bin1, 
                              y = MSE, 
                              colour = method)) + 
  geom_point(size = 0.8) +
  geom_line(linewidth = 0.3) + 
  # geom_vline(xintercept = 15, linetype="solid", 
  #            color = "black", linewidth=0.8, alpha = 0.5) +
  scale_color_manual(values=c('#ff7f00', '#999999', 
                              '#a65628', '#f781bf')) +
  ylab("MSE") +
  xlab(expression(b[1])) + 
  theme(aspect.ratio = 0.75,
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 7),
        axis.title.x = element_text(size = 7),
        axis.title.y = element_text(size = 7, angle = 90),
        axis.ticks.x = element_line(linewidth = 0.5),
        axis.ticks.y = element_line(linewidth = 0.5),
        legend.position = "none") +
  inset_element(p = pbmc_layouts, 
                left = 1.55, 
                bottom = 0.35, 
                right = 0, 
                top = 1)

@@

```{r}
#| fig-cap: "MSE from UMAP with diffferent (hyper-)parameter choices and tSNE with default (hyper-)parameter setting applied to PBMC3k dataset for different number of bins along the x-axis ($b_1$). Which is the most reasonable representation of the structures present in the PBMC3k data? The residual plot have a steep slope at the beginning, indicating that a smaller $b_1$ causes a larger amount of error. Then, the slope gradually declines or level off, indicating that a higher number of $b_1$ generates a smaller error. It was observed that for every $b_1$, the lowest error occurred with tSNE using the perplexity: $30$."
#| label: fig-pbmc-abserror
#| fig-pos: H
#| out-width: 70%

### 2024-08-19 paper.tex
Updated Figure 15

\caption{\label{fig-pbmc-abserror}Absolute error from UMAP and tSNE
applied to PBMC3k dataset with diffferent (hyper-)parameter choices.
What is the best (hyper-)parameter choice to create the model? The
residual plot have a steep slope at the beginning, indicating that a
smaller number of non-empty bins causes a larger amount of error. Then,
the slope gradually declines or level off, indicating that a higher
number of non-empty bins generates a smaller error. Using the elbow
method, it was observed that when the number of non-empty bins is set to
\(136\), the lowest error occurred with the parameters perplexity:
\(30\).}

>>>>

\caption{\label{fig-pbmc-abserror}MSE from UMAP with diffferent
(hyper-)parameter choices and tSNE with default (hyper-)parameter
setting applied to PBMC3k dataset for different number of bins along the
x-axis (\(b_1\)). Which is the most reasonable representation of the
structures present in the PBMC3k data? The residual plot have a steep
slope at the beginning, indicating that a smaller \(b_1\) causes a
larger amount of error. Then, the slope gradually declines or level off,
indicating that a higher number of \(b_1\) generates a smaller error. It
was observed that for every \(b_1\), the lowest error occurred with tSNE
using the perplexity: \(30\).}
